---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  warning = FALSE,
  message = FALSE
)
```

# bullseye

<!-- badges: start -->
<!-- badges: end -->

The goal of bullseye is to provide a tidy data structure and visualisations for multiple or grouped  variable correlations, 
    general association measures and other pairwise scores  suitable for numerical, ordinal and nominal variables.

## Installation

You can install the development version of bullseye from [GitHub](https://github.com/) with:

``` {r}
# still in progress
# install.packages("devtools")
# devtools::install_github("cbhurley/bullseye")
```

## Build a `pairwise` data structure

```{r}
library(bullseye)
irisc <- pairwise_scores(iris, by = "Species") 
irisc
```

This calculates correlations for every level of species in the data.

For other methods for calculating pairwise variable scores see the  vignette "Calculating pairwise scores using bullseye".


## Visualise the correlations

```{r vis1, fig.height=5.5, fig.width=5}
plot_pairwise(irisc)
```

The glyph shows the overall or ungrouped correlation in the center, and the grouped correlations in the wedges. These are setosa, versicolor, virginica starting from the top and going anti-clockwise. Use 

```{r , eval=F}
plot_pairwise(irisc, interactive=TRUE)
```

to see the groups represented by the wedges.  The ungrouped Pearson correlation of Sepal.Width with the other three variables is negative, but the correlations for each of the Species is positive. 

A second visualisation is offered also:

```{r vis2, fig.height=5.5, fig.width=5}
plot_pairwise_linear(irisc)
```

For more visualisation examples  see the  vignette "Visualising pairwise scores using bullseye".
