---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

# bullseye

<!-- badges: start -->
<!-- badges: end -->

The goal of bullseye is to provide a tidy data structure and visualisations for multiple or grouped  variable correlations, 
    general association measures and other pairwise scores  suitable for numerical, ordinal and nominal variables.

## Installation

You can install the development version of bullseye from [GitHub](https://github.com/) with:

``` {r}
# still in progress
# install.packages("devtools")
# devtools::install_github("cbhurley/bullseye")
```

## Build a `pairwise` data structure

```{r, eval=F}
irisc <- pair_scores(iris, by = "Species") 
irisc
```

This calculates correlations for every level of species in the data.


## Visualise the correlations

```{r, eval=F}
plot_pairwise(irisc)
```

